\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-04-19 doc/sw-system.tex
Fixed #131

\new{Because we include a training set of images in amongst the survey ``test''
images in this way, \sw can be viewed as a {\it supervised learning} system.
Consequently, we should expect to find lens candidates that look like the sims
that we put in the training set, and so the design of this training set is
quite important. The training images must look realistic, in two ways. First,
apart from the simulated lenses they contain, they must look like all the
other survey images -- otherwise they could be identified as training images
without the volunteer learning anything about what lenses look like. For this
reason, we make the sims by {\it adding in} simulated lensed features to real
images drawn from the survey. It also means that each dataset presented for
inspection in the \sw interface needs to come with its own training set.}

\new{Second, the lensed features themselves must be realistic; if they did not
resemble real lenses, we could not expect to find real lenses in the test
images. The details of how we generated the sims for the CFHTLS project are
given in \PaperTwo, where we also carry out performance tests relative to real
lens candidates in the survey fields. Here, we merely note the following
aspects to place this paper's results in context.}

\new{We start by selecting massive galaxies and groups in the CFHTLS catalogs,
by color, brightness and photometric redshift. We then assign plausible mass
distributions to these ``potential lenses:'' a singular isothermal ellipsoid
model for each massive galaxy, and an additional NFW profile dark matter halo
for groups and clusters. The mass distributions are centred, elongated and
aligned with the measured optical properties of the galaxies present. We then
draw a source from a plausible background population, of either galaxies (for
the group or galaxy lenses) or quasars (for the galaxy-scale lenses). The
source redshift, luminosity and size distributions are all chosen to match
those observed, while the source abundance is artificially increased such that
each potential lens has a high probability of having a source occur at a
position that will lead to it being  highly magnified. Source galaxies are
given plausible surface brightness profiles and ellipticities. We then compute
the predicted multiple images of the source using the \gravlens ray-tracing
code \citep{Keeton2000}.}

At this point, the distribution of lens and source properties are quite
realistic. We then chose to make cuts of the simulated lens pool, in order to
reject sims that were likely to be diﬃcult to spot. While this ensured the
educational value of the sims when feedback was given, it meant that we should
only expect to detect real lens candidates that also pass these cuts. Any real
lens candidates we ﬁnd with properties outside the ranges of the sims, may in
some sense be considered bonuses, a product of the {\it imagination} of the
crowd in extrapolating their understanding of gravitational lenses into a regime
where they have not seen any before. In the CFHTLS project, the most important
selection cuts were to only keep sims with a)  second brightest image brighter
that $i = 23$ and b) total combined image magnitude brighter than 19--20
(\PaperTwo, Table~1).} The combination of these two cuts results in a sample of
visible lensed image configurations.

@@

The CFHTLS $g$, $r$ and $i$-band images have the greatest depth and best image
quality, and we chose to focus on this set (although the $u$ and $z$-band images
were also available for perusal in Talk). \new{We made colour composite PNG
format images following the prescription of \citet{LuptonEtal2004} \citep[with
extensions by][and some particular choices of our own]{WherryEtal2004}, using
the \humvi software.\footnote{The \humvi colour image composition code used in
this work is open source and available from
\texttt{http://github.com/drphilmarshall/HumVI}} Specifically, we first rescaled
the pixel values of each image, in the notation of \citeauthor{LuptonEtal2004},
into flux units (``picomaggies,''), via the image AB zeropoint $m_0$ and the
pixel value calibration factor $f$  given by $\log_{10}{f} = 0.4(30.0 - m_0)$.
We then multiplied these calibrated images by further aesthetic ``scales''
$s_{i,r,g}$ before computing the total intensity image $I$ and applying an
arcsinh stretch. (The scales are re-normalized to sum to one on input.) Thus,
the red ($R$), green ($G$) and blue ($B$) channel images correspond to the
CFHTLS $i$, $r$ and $g$-band images in the following way:
\bea
I & = & (i \cdot s_i + r \cdot s_r + g \cdot s_g), \notag \\
R & = & i \cdot s_i \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}, \notag \\
G & = & r \cdot s_r \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}, \notag \\
B & = & g \cdot s_g \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}. \notag \\
\eea
We chose to allow the composite image formed from these channel images to
saturate to white: any pixels in any of the channel images lying outside the
range 0 to 1 was assigned the value 0 or 1 appropriately. This was not the
recommendation of \citet{LuptonEtal2004}, but we found it to still give very
informative but also familiar-looking astronomical images.}

\new{The non-linearity parameters $Q$ and $\alpha$ control the brightness and
contrast of the images. We first tuned $\alpha$ (which acts as an additional
scale factor) until the background noise was just visible. Then we tuned the
colour scales $s$ to find a balance between exposing the low surface
brightness blue features (important for lens spotting!), and having the noise
appear to have equal red, green and blue components. Non-linearity sets in at
about $1/(Q \alpha)$ in the  scaled intensity image: finally, tuning $Q$ at
fixed $\alpha$ determines the appearance of bright galaxies, which we need to
be suppressed enough to allow the low surface brightness features to show
through, but not so much that they no longer looked like massive galaxies.}

>>>>

Because we include a training set of images in amongst the survey ``test''
images, \sw can be viewed as a {\it supervised learning} system.
Consequently, we should expect to find lens candidates that look like the sims
that we put in the training set, and so the design of this training set is
quite important. The training images must look realistic, in two ways. First,
apart from the simulated lenses they contain, they must look like all the
other survey images -- otherwise they could be identified as training images
without the volunteer learning anything about what lenses look like. For this
reason, we make the sims by {\it adding in} simulated lensed features to real
images drawn from the survey. It also means that each dataset presented for
inspection in the \sw interface needs to come with its own training set.

Second, the lensed features themselves must be realistic; if they did not
resemble real lenses, we could not expect to find real lenses in the test
images. The details of how we generated the sims for the CFHTLS project are
given in \PaperTwo, where we also carry out performance tests relative to real
lens candidates in the survey fields. Here, we merely note the following
aspects to place this paper's results in context.

We start by selecting massive galaxies and groups in the CFHTLS catalogs,
by color, brightness and photometric redshift. We then assign plausible mass
distributions to these ``potential lenses:'' a singular isothermal ellipsoid
model for each massive galaxy, and an additional NFW profile dark matter halo
for groups and clusters. The mass distributions are centred, elongated and
aligned with the measured optical properties of the galaxies present. We then
draw a source from a plausible background population, of either galaxies (for
the group or galaxy lenses) or quasars (for the galaxy-scale lenses). The
source redshift, luminosity and size distributions are all chosen to match
those observed, while the source abundance is artificially increased such that
each potential lens has a high probability of having a source occur at a
position that will lead to it being  highly magnified. Source galaxies are
given plausible surface brightness profiles and ellipticities. We then compute
the predicted multiple images of the source using the \gravlens ray-tracing
code \citep{Keeton2000}.

At this point, the distribution of lens and source properties are quite
realistic. We then chose to make cuts of the simulated lens pool, in order to
reject sims that were likely to be diﬃcult to spot. While this ensured the
educational value of the sims when feedback was given, it meant that we should
only expect to detect real lens candidates that also pass these cuts. Any real
lens candidates we ﬁnd with properties outside the ranges of the sims, may in
some sense be considered bonuses, a product of the {\it imagination} of the
crowd in extrapolating their understanding of gravitational lenses into a regime
where they have not seen any before. In the CFHTLS project, the most important
selection cuts were to only keep sims with a)  second brightest image brighter
that $i = 23$ and b) total combined image magnitude brighter than 19--20
(\PaperTwo, Table~1). The combination of these two cuts results in a sample of
visible lensed image configurations.

@@

The CFHTLS $g$, $r$ and $i$-band images have the greatest depth and best image
quality, and we chose to focus on this set (although the $u$ and $z$-band images
were also available for perusal in Talk). We made colour composite PNG
format images following the prescription of \citet{LuptonEtal2004} \citep[with
extensions by][and some particular choices of our own]{WherryEtal2004}, using
the \humvi software.\footnote{The \humvi colour image composition code used in
this work is open source and available from
\texttt{http://github.com/drphilmarshall/HumVI}} Specifically, we first rescaled
the pixel values of each image, in the notation of \citeauthor{LuptonEtal2004},
into flux units (``picomaggies,''), via the image AB zeropoint $m_0$ and the
pixel value calibration factor $f$  given by $\log_{10}{f} = 0.4(30.0 - m_0)$.
We then multiplied these calibrated images by further aesthetic ``scales''
$s_{i,r,g}$ before computing the total intensity image $I$ and applying an
arcsinh stretch. (The scales are re-normalized to sum to one on input.) Thus,
the red ($R$), green ($G$) and blue ($B$) channel images correspond to the
CFHTLS $i$, $r$ and $g$-band images in the following way:
\bea
I & = & (i \cdot s_i + r \cdot s_r + g \cdot s_g), \notag \\
R & = & i \cdot s_i \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}, \notag \\
G & = & r \cdot s_r \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}, \notag \\
B & = & g \cdot s_g \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}. \notag \\
\eea
We chose to allow the composite image formed from these channel images to
saturate to white: any pixels in any of the channel images lying outside the
range 0 to 1 was assigned the value 0 or 1 appropriately. This was not the
recommendation of \citet{LuptonEtal2004}, but we found it to still give very
informative but also familiar-looking astronomical images.

The non-linearity parameters $Q$ and $\alpha$ control the brightness and
contrast of the images. We first tuned $\alpha$ (which acts as an additional
scale factor) until the background noise was just visible. Then we tuned the
colour scales $s$ to find a balance between exposing the low surface
brightness blue features (important for lens spotting!), and having the noise
appear to have equal red, green and blue components. Non-linearity sets in at
about $1/(Q \alpha)$ in the  scaled intensity image: finally, tuning $Q$ at
fixed $\alpha$ determines the appearance of bright galaxies, which we need to
be suppressed enough to allow the low surface brightness features to show
through, but not so much that they no longer looked like massive galaxies.
